Expression QTL hotspot dissection with mediation analysis and pleiotropy testing

### 2019-07-31 analysis/paper/hotspots-paper.Rmd
updated paper text

```{r}
if (!file.exists(here("analysis", "data", "derived_data", "expr.rds"))){
  download.file("https://datadryad.org/bitstream/handle/10255/dryad.166774/Attie_DO378_eQTL_viewer_v1.Rdata?sequence=2",
              destfile = here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata")
              )
  load(here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata"))
  saveRDS(dataset.islet.rnaseq$lod.peaks, here("analysis", "data", "derived_data", "lod_peaks.rds"))
  saveRDS(dataset.islet.rnaseq$annots, here("analysis", "data", "derived_data", "annots.rds"))
  saveRDS(dataset.islet.rnaseq$expr, here("analysis", "data", "derived_data", "expr.rds"))
}
```

@@

>>>>

```{r}
if (!file.exists(here("analysis", "data", "derived_data", "expr.rds"))){
  download.file("https://datadryad.org/bitstream/handle/10255/dryad.166774/Attie_DO378_eQTL_viewer_v1.Rdata?sequence=2",
              destfile = here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata")
              )
  load(here("analysis", "data", "raw_data", "Attie_DO378_eQTL_viewer_v1.Rdata"))
  saveRDS(dataset.islet.rnaseq$lod.peaks, here("analysis", "data", "derived_data", "lod_peaks.rds"))
  saveRDS(dataset.islet.rnaseq$annots, here("analysis", "data", "derived_data", "annots.rds"))
  saveRDS(dataset.islet.rnaseq$expr, here("analysis", "data", "derived_data", "expr.rds"))
  saveRDS(K, here("analysis", "data", "derived_data", "kinship.rds"))
  saveRDS(map, here("analysis", "data", "derived_data", "map.rds"))
  saveRDS(genoprobs, here("analysis", "data", "derived_data", "genoprobs.rds"))
  saveRDS(dataset.islet.rnaseq$covar, here("analysis", "data", "derived_data", "covar.rds"))
}
```

@@

An unanticipated observation is that some points have mediation LOD differences of nearly 10 
and pleiotropy test statistics of nearly 10. With such large pleiotropy test statistics, one 
has reasonable evidence each point representing two separate QTL. It seems biologically 
unlikely that a nonlocal trait and a mediating local trait would not share a pleiotropic QTL.

In looking closer at these six points, we find that five of the six have *Tmem189* as the 
local trait. This raises questions about the *Tmem189* trait. One known issue with the pleiotropy test is its sensitivity to start and end points for the two-dimensional scan. If the scan interval is too small, the test can return misleading results. If *Tmem189* were to have a very broad univariate peak, such that the peak extended beyond the scan interval, then we might observe misleading results in tests involving *Tmem189*.

```{r scan1}
# read oddball points from interactive plot
tr <- c("Tmem189",
        "Myo15b",
        "Sephs2",
        "Bcmo1",
        "Tmprss4",
        "Cacnb3",
        "Gm12929",
        "Eya2"
        )
tr_tib <- hot_annot %>%
  filter(symbol %in% tr) %>%
  arrange(annot.id)
K <- readRDS(here("analysis", "data", "derived_data", "kinship.rds"))
map <- readRDS(here("analysis", "data", "derived_data", "map.rds"))
covar <- readRDS(here("analysis", "data", "derived_data", "covar.rds"))
aprobs <- readRDS(here("analysis", "data", "derived_data", "genoprobs.rds"))
expr_sm <- expr[ , colnames(expr) %in% tr_tib$annot.id, drop = FALSE]
s1 <- qtl2::scan1(genoprobs = aprobs,
            pheno = expr_sm,
            kinship = K,
            addcovar = covar, 
            cores = 0
)
# zoom in on Chr2 after 150 Mb
# get length of map$`1`
purrr::map(.x = 1:8, .f = function(x) qtl2::plot_scan1(s1, map, lodcolumn = x, chr = 2, main = paste0(tr_tib[x, ])))
qtl2::find_peaks(scan1_output = s1, map = map, threshold = 7) %>% 
  as_tibble() %>%
  filter(chr == 2) %>%
  left_join(hot_annot, by = c("lodcolumn" = "annot.id"))
```
